
---
title: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Statistical Computing - Exercises 05 - Variable Selection

In these exercises, you will implement a regression variable selection
procedure called forward selection. You will apply your variable
selection procedure to the Ames, Iowa housing dataset, which can be
read into R with (after you've installed the modeldata package):

```{r}
#install.packages("modeldata")
data("ames", package = "modeldata")
```

Forward selection starts with a model with only an intercept, and then
picks the covariate that, when added to the model, minimizes an
*information criterion*, such as the Akaike Information Criterion
(AIC). After that covariate is added, it attempts to add the covariate
from the remaining set of covariates that when added to the model,
makes the information criterion smallest. This process is continued
until none of the remaining covariates decrease the information
criterion.

1. Look up the definition of the AIC (Wikipedia is fine), and write a
function that takes in an object of class `lm` (the result of running
the `lm` function to fit a linear model), and returns the AIC value.
There is a function in R called `logLik` that might be helpful.

2. Write a function that takes in a string indicating the response
variable, and a vector of strings indicating covariate values, and
returns a string containing a model formula, such as
\verb!"Sale_Price ~ Gr_Liv_Area + Total_Bsmt_SF"!

3. Write a function that takes in a string response, two vectors of
string covariates, a dataset, and returns the covariate from the
second set that decreases the AIC most when added to a model
containing the first set of covariates. If none decrease the AIC, the
function should return NULL or a length-zero vector.

4. Use these functions to write a function implementing the forward
selection algorithm on the Ames housing data, using both Sale_Price
and log Sale_Price as the response. Your forward selection function
should take in a response string, a character vector of candidate
covariates (in case you don't want to test all the covariates, or want
to try some interactions), a data frame, and return the vector of
selected covariates. Your function should check whether the candidate
covariates are available in the data frame, and return an error with
an informative error message if not. 

